Towards human SuperEEG

### 2018-10-05 paper/main.tex
working on results; WIP

@@

\begin{align}
\hat{N}_{s}(x,y) & = { \sum_{i = 1}^{| R_{s}|}\sum_{j=1}^{i-1} W(x,i) \cdot W(y,j)\cdot \mathrm{z}(\bar{C}_{s}(i,j))}\label{eqn:subj_corrmat_num}\\
 \hat{D}_{s}(x,y) & = \sum_{i = 1}^{| R_{s}|}\sum_{j=1}^{i-1} W(x,i)
                    \cdot W(y,j). \label{eqn:subj_corrmat_den}\\
 \hat{C}_s &= \mathrm{r}\left( \frac{\hat{N}_{s}}{\hat{D}_{s}} \right)
             \label{eqn:subj_corrmat}.
\end{align}
After estimating the numerator ($\hat{N}_{s}$) and denominator
($\hat{D}_{s}$) placeholders for each $\hat{C}_{s}$, we aggregate these
estimates across patients to obtain a single expected full-brain
correlation matrix ($\hat{K}$; Fig.~\ref{fig:methods}D):
\begin{align}
 \hat{K} = \mathrm{r} \left(  \frac{\sum  \hat{N}_{s}}{\sum \hat{D}_{s}}\right).\label{eqn:correlation_matrix}
\end{align}
Intuitively, the numerators capture the general structures of the
patient-specific estimates of full-brain correlations, and the
denominators account for which locations were near the implanted
electrodes in each patient.  To obtain $\hat{K}$, we compute a
weighted average across the estimated patient-specific full-brain
correlation matrices, where patients with observed electrodes near a
particular set of locations in $\hat{K}$ contribute more to the
estimate.

@@

\begin{align}
Y_{s,k,\beta} &= ((\hat{K}_{\beta_s,\alpha_s}\cdot\hat{K}_{\alpha_s,\alpha_s}^{-1})\cdot Y_{s,k,\alpha_s}^\mathrm{T})^\mathrm{T}.\label{eqn:reconstruction}
\end{align}
This equation is the foundation of the SuperEEG algorithm.  Whereas we
observe recordings only at the locations in $\alpha_s$,
Equation~\ref{eqn:reconstruction} allows us to estimate the recordings
at all locations in $\beta_s$, which we can define \textit{a priori}
to include any locations we wish, throughout the brain.  This yields
estimates of the time-varying voltages at \textit{every} location in
$\bar{R}$, provided that we define $\bar{R}$ in advance to include the
union of all of the locations in $\alpha_s$ and all of the locations
at which we wish to estimate recordings (i.e., a timeseries of
voltages).

@@

\section*{Results}
To test the accuracy with which the SuperEEG algorithm reconstructs
activity throughout the brain, we held out each electrode from the
full dataset in turn and treated it as unobserved. We then asked: how
closely did each of the SuperEEG-estimated recordings at those
electrodes match the observed recordings from those electrodes?

We first examined individual held-out raw LFP traces and their
associated SuperEEG-derived reconstructions.  Figure~\ref{fig:recon}A
displays the observed LFP from the red electrode in
Figure~\ref{fig:methods}A (red), and its associated reconstruction
(blue), during a 5~s time window during one of the example patient's
six recording sessions.  The two traces match closely
($r = XXX, p = XXX$).
%Where the traces do not align, the predicted
%trace tends to be rounded towards zero, reflecting the zero-mean prior
%implicit in Equation~\ref{eqn:reconstruction} (paired $t$-test
%comparing absolute values of actual versus predicted voltages:
%$t(XXX) = XXX, p = XXX$).
Figure~\ref{fig:recon}B displays a
two-dimensional histogram of the actual versus reconstructed voltages
for the entire 14.2 total hours of recordings from the electrode
(correlation: $r = XXX, p = XXX$).
%Like in the shorter example trace
%shown in panel A, the reconstructed voltages throughout the full
%recording tended to be smaller in magnitude than the original
%recording ($t(XXX) = XXX, p = XXX$).
Having confirmed that the SuperEEG algorithm recovered the recordings
from this electrode well, we next sought to quantify the algorithm's
performance across the full dataset.

%sample trace & reconstruction, several frames of activity snapshot?
\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{figs/recon}
  \caption{\textbf{Observed and reconstructed LFP from a single
      electrode.} \textbf{A. Example LFP.}  A 5~s recording from the
    red electrode in Figure~\ref{fig:methods}A is displayed in red,
    and the reconstructed LFP during the same time window is shown in
    blue.  \textbf{B. Observed versus reconstructed voltages over
      14.2~hours.}  The two-dimensional histogram reflects the
    relation between distributions of observed versus reconstructed
    voltages from one patient, across the 14.2 hours of recorded data
    collected over 6 recording sessions.  The correlation reported in
    the panel is between the observed and reconstructed voltages.
    Both panels: all voltages are represented in standard deviation units
   (computed within session).}
  \label{fig:recon}
\end{figure}

Holding out each electrode from each patient in turn, we computed the
average correlation (across recording sessions) between the Super
EEG-reconstructed voltage traces and the observed voltage traces from
that electrode.  For each reconstruction, we estimated the full-brain
correlation matrix using every \textit{other} patient's data
(i.e., every patient except the one who contributed the
to-be-reconstructed electrode data).  In our analyses, we then
substituted the average correlation matrix computed after excluding
patient $s$'s data for $\hat{K}$ in Equations~\ref{eqn:Kba} and
\ref{eqn:Kaa}.  This step ensured that the data we were reconstructing
could not also be used to estimate the between-location correlations
that drove the reconstructions via Equation~\ref{eqn:reconstruction}
(otherwise the analysis would be circular).

We obtained a single correlation coefficient for each electrode
location in $\bar{R}$, reflecting how well the SuperEEG algorithm was
able to recover the recording at that location by incorporating data
across patients (Across shown in black, see
Fig.~\ref{fig:corrmap}A). We also reconstructed activity for each
electrode using a model trained on the remaining electrodes from only
that patient, to account for reconstruction accuracy attributed to
volume conductance alone (Within shown in gray, see
Fig.~\ref{fig:corrmap}A). For the first dataset, we compared these two
distributions of correlation coefficients (paired $t$-test between
$z$-transformed mean correlation coefficients by patient:
$t(66) = 9.64, p < 10\textsuperscript{-10}$).  We repeated this
analysis on a similar dataset Fig.~\ref{fig:corrmap}C) with similar
results (paired $t$-test between $z$-transformed mean correlation
coefficients by patient: $t(23) = 6.93, p <
10\textsuperscript{-5}$). This is an especially conservative test,
given that the SuperEEG reconstructions exclude (from the correlation
matrix estimates) all data from the patient whose data is being
reconstructed.  Furthermore, we also replicated this finding for each
independent experiment within dataset 2 (Fig. S3 (paired $t$-test
between $z$-transformed mean correlation coefficients by patient for
experiment 1: $t(23) = 6.23, p < 10\textsuperscript{-5}$ and
experiment 2: $t(23) = 6.62, p < 10\textsuperscript{-5}$).  That the
SuperEEG-derived correlations were reliably stronger than these
correlations obtained using a volume conductance null model is
exciting for two reasons.  First, it implies that distant electrodes
provide additional predictive power to the data reconstructions beyond
the information contained in nearby electrodes.  Second, it implies
that the spatial correlations driving the SuperEEG algorithm are, to
some extent, shared across people.

%correlations by brain area, distribution of correlations
\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{figs/corrmap}
  \caption{\textbf{Reconstruction quality.}  \textbf{A. \& C.  Distributions
      of correlation coefficients.}  Across all electrodes from all
    patients in the labeled dataset, the panel displays the distribution of correlations between the observed and reconstructed LFP data using models trained on data from all other patients (Across, in black) and all other electrodes from the same patient (Within, in gray). 
    \textbf{B. \& D. Correlation maps.}  The glass brain maps display the
    average correlation between the observed LFP data and the across-subjects model reconstructed data by location, for each labeled experiment.}
  \label{fig:corrmap}
\end{figure}

We were interested in the task specific contributions to the reconstruction accuracy.  Each patient in the the second dataset participated in two free recall experiments.  We ran similar analyses for both experiments and found that activity was best reconstructed when limiting the training data to within task, as opposed to across task or incorporating data from both tasks (Fig. S1 (mean reconstruction accuracy incorporating data within task: 0.55, across task: 0.37, all tasks: .50)). Although reconstruction accuracy in the across task analysis was still better than the volume conductance model alone (paired $t$-test
between $z$-transformed mean correlation coefficients by patient: $t(47) = 5.65, p < 10\textsuperscript{-5}$), these results suggests that having a common tasks for patients may yield better reconstruction accuracy.

>>>>

\newcommand{\taskrecon}{S1}
\newcommand{\per_exp_taskrecon}{S2}
\newcommand{\per_exp_corrmaps}{S3}

@@

\begin{align}
\hat{N}_{s}(x,y) & = { \sum_{i = 1}^{| R_{s}|}\sum_{j=1}^{i-1} W(x,i) \cdot W(y,j)\cdot \mathrm{z}(\bar{C}_{s}(i,j))}\label{eqn:subj_corrmat_num}\\
 \hat{D}_{s}(x,y) & = \sum_{i = 1}^{| R_{s}|}\sum_{j=1}^{i-1} W(x,i)
                    \cdot W(y,j). \label{eqn:subj_corrmat_den}\\
 \hat{C}_s &= \mathrm{r}\left( \frac{\hat{N}_{s}}{\hat{D}_{s}} \right)
             \label{eqn:subj_corrmat}.
\end{align}
After estimating the numerator ($\hat{N}_{s}$) and denominator
($\hat{D}_{s}$) placeholders for each $\hat{C}_{s}$, we aggregate these
estimates across patients to obtain a single expected full-brain
correlation matrix ($\hat{K}$; Fig.~\ref{fig:methods}D):
\begin{align}
 \hat{K} = \mathrm{r} \left(  \frac{\sum  \hat{N}_{s}}{\sum \hat{D}_{s}}\right).\label{eqn:corrmat}
\end{align}
Intuitively, the numerators capture the general structures of the
patient-specific estimates of full-brain correlations, and the
denominators account for which locations were near the implanted
electrodes in each patient.  To obtain $\hat{K}$, we compute a
weighted average across the estimated patient-specific full-brain
correlation matrices, where patients with observed electrodes near a
particular set of locations in $\hat{K}$ contribute more to the
estimate.

@@

\begin{align}
\hat{Y}_{s,k,\beta_s} &= ((\hat{K}_{\beta_s,\alpha_s}\cdot\hat{K}_{\alpha_s,\alpha_s}^{-1})\cdot Y_{s,k,\alpha_s}^\mathrm{T})^\mathrm{T}.\label{eqn:reconstruction}
\end{align}
This equation is the foundation of the SuperEEG algorithm.  Whereas we
observe recordings only at the locations in $\alpha_s$,
Equation~\ref{eqn:reconstruction} allows us to estimate the recordings
at all locations in $\beta_s$, which we can define \textit{a priori}
to include any locations we wish, throughout the brain.  This yields
estimates of the time-varying voltages at \textit{every} location in
$\bar{R}$, provided that we define $\bar{R}$ in advance to include the
union of all of the locations in $\alpha_s$ and all of the locations
at which we wish to estimate recordings (i.e., a timeseries of
voltages).

@@

\section*{Results}
We used a cross-validation approach to test the accuracy with which
the SuperEEG algorithm reconstructs activity throughout the brain.
For each patient in turn, we estimated full-brain correlation matrices
(Eqn.~\ref{eqn:corrmat}) using data from all of the \textit{other}
patients. This step ensured that the data we were reconstructing could
not also be used to estimate the between-location correlations that
drove the reconstructions via Equation~\ref{eqn:reconstruction}
(otherwise the analysis would be circular).  For that held-out
patient, for each of their electrodes in turn, we used
Equation~\ref{eqn:reconstruction} to reconstruct activity at the
held-out electrode location, using the correlation matrix trained on
all other patients' data at $\hat{K}$, and using activity recorded
from the other electrodes from the held-out patient as
$Y_{s, k, \alpha_s}$.  We then asked: how closely did each of the
SuperEEG-estimated recordings at those electrodes match the observed
recordings from those electrodes (i.e., how closely did the estimated
$\hat{Y}_{s, k, \beta_s}$ match the observed $Y_{s, k, \beta_s}$?).

To illustrate our approach, we first examine and individual held-out
raw LFP trace and its associated SuperEEG-derived reconstruction.
Figure~\ref{fig:recon}A displays the observed LFP from the red
electrode in Figure~\ref{fig:methods}A (red), and its associated
reconstruction (blue), during a 5~s time window during one of the
example patient's six recording sessions.  The two traces match
closely ($r = XXX, p = XXX$).
%Where the traces do not align, the predicted
%trace tends to be rounded towards zero, reflecting the zero-mean prior
%implicit in Equation~\ref{eqn:reconstruction} (paired $t$-test
%comparing absolute values of actual versus predicted voltages:
%$t(XXX) = XXX, p = XXX$).
Figure~\ref{fig:recon}B displays a
two-dimensional histogram of the actual versus reconstructed voltages
for the entire 14.2 total hours of recordings from the example electrode
(correlation: $r = XXX, p = XXX$).
%Like in the shorter example trace
%shown in panel A, the reconstructed voltages throughout the full
%recording tended to be smaller in magnitude than the original
%recording ($t(XXX) = XXX, p = XXX$).
This example confirms that the SuperEEG algorithm recovers the recordings
from this single electrode well.  Next, we quantify the algorithm's
performance across the full dataset.

%sample trace & reconstruction, several frames of activity snapshot?
\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{figs/recon}
  \caption{\textbf{Observed and reconstructed LFP from a single
      electrode.} \textbf{A. Example LFP.}  A 5~s recording from the
    red electrode in Figure~\ref{fig:methods}A is displayed in red,
    and the reconstructed LFP during the same time window is shown in
    blue.  \textbf{B. Observed versus reconstructed LFP over
      14.2~hours.}  The two-dimensional histogram reflects the
    relation between distributions of observed versus reconstructed
    voltages from one patient, across the 14.2 hours of recorded data
    collected over 6 recording sessions.  The correlation reported in
    the panel is between the observed and reconstructed voltages.
    Both panels: all voltages are represented in standard deviation units
   (computed within session).}
  \label{fig:recon}
\end{figure}

For each held-out electrode, from each held-out patient in turn, we we
computed the average correlation (across recording sessions) between
the SuperEEG-reconstructed voltage traces and the observed voltage
traces from that electrode.  For this analysis we set $\bar{R}$ to be
the union of all electrode locations across all patients.  This
yielded a single correlation coefficient for each electrode location
in $\bar{R}$, reflecting how well the SuperEEG algorithm was able to
recover the recording at that location by incorporating data across
patients (black histogram in Fig.~\ref{fig:corrmap}A, map in
Fig.~\ref{fig:corrmap}C).  The observed distribution of correlations
was centered well above zero (mean: XXX; $t$-test comparing mean of
distribution of $z$-transformed per-electrode correlation coefficients
to 0: $t(XXX) = XXX, p = XXX$), indicating that the SuperEEG algorithm
recovers held-out activity patterns substantially better than random
guessing.

As a stricter benchmark, we compared the quality of these
across-participant reconstructions (i.e., computed using a correlation
model derived from other patients' data) to reconstructions generated
using a correlation model trained using the in-patient's data.  In
other words, for this within-patient benchmark analysis we estimated
$\hat{C}_{s}$ (Eqn.~\ref{eqn:subj_corrmat}) for each patient in turn,
using recordings from all of that patient's electrodes except at the
location we were reconstructing.  These within-patient reconstructions
serve as an estimate of how well data from all of the other electrodes
from a single patient may be used to estimate held-out data.  This
allows us to ask how much information about the activity at a given
electrode might be inferred through (a) volume conductance or other
sources of ``leakage'' from activity patterns measured from the
patient's other electrodes and (b) across-electrode correlations
learned from that single patient.  As shown in
Figure~\ref{fig:corrmap}A (gray histogram), the distribution of
within-patient correlations was centered well above zero (mean: XXX;
$t$-test comparing mean of distribution of $z$-transformed
per-electrode correlation coefficients to 0: $t(XXX) = XXX, p = XXX$).
However, the across-patient correlations were substantially higher
($t$-test comparing average $z$-transformed within versus across
patient electrode correlations: $t(XXX) = XXX, p = XXX$).  This is an
especially conservative test, given that the across-patient SuperEEG
reconstructions exclude (from the correlation matrix estimates) all
data from the patient whose data is being reconstructed.  We repeated
each of these analyses on a second independent datset and found
similar results (Fig.~\ref{corrmap}B, D; within versus across
reconstruction accuracy: $t(23) = 6.93, p < 10^{-5}$). Our finding,
that when reconstructing held-out data from a given patient
correlation models derived from \textit{other} patient's data yield
higher reconstruction accuracy than correlation models derived from
that patient, has two important implications.  First, it implies that
distant electrodes provide additional predictive power to the data
reconstructions beynd the information contained solely in nearby
electrodes.  (This follows from the fact that each patient's
electrodes are implanted in a unique set of locations, so for any
given electrode the closest electrodes in the full dataset are likely
to come from the same patient.)  Second, it implies that the spatial
correlations derived from the SuperEEG algorithm are, to some extent,
similar across people.

%correlations by brain area, distribution of correlations
\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{figs/corrmap}
  \caption{\textbf{Reconstruction quality across all electrodes in two
      ECoG datasets.}  \textbf{A. Distributions of correlations
      between observed versus reconstructed activity by electrode, for
      Dataset 1.}  The across-patient distribution (black) reflects
    reconstruction performance using a correlation model trained on
    all but one patient, and then applied to the held-out patient's
    data.  The within-patient distribution (gray) reflects performance using
    a correlation model trained on the same patient who contributed
    the to-be-reconstructed electrode.  \textbf{B. Distributions of
      correlations for Dataset 2.}  This panel is in the same format
    as Panel A, but reflects results obtained from Dataset 2.
    \textbf{C.--D.  Reconstruction performance by location.} Each dot
    reflects the location of a single implanted electrode from Dataset
    1 (Panel C) or Dataset 2 (Panel D).  The dot colors denote the
    average across-session correlation, using the across-patient
    correlation model, between the observed and
    reconstructed activity at the given electrode location.}
  \label{fig:corrmap}
\end{figure}

The recordings we analyzed from Dataset 1 comprised data collected as
the patients performed a variety of (essentially uncontrolled) tasks
throughout each day's recording session.  That we observed reliable
reconstruction accuracy across patients suggests that the spatial
correlations derived from the SuperEEG algorithm are, to some extent,
similar across tasks.  We tested this finding more explicitly using
Dataset 2.  In Dataset 2, the recordings were limited to times when
each patient was participating in each of two experiments (Experiment
1, a random-word list free recall task, and Experiment 2, a
categorized list free recall task).  We wondered whether a correlation
model trained using data only from one experiment might yield good
predictions for data from the other experiment.  Further, we wondered
about the extent to which it might be benefitial or harmful to combine
data across tasks.

\textbf{JRM STOPPED HERE}

We were interested in the task specific contributions to the
reconstruction accuracy.  Each patient in the the second dataset
participated in two free recall experiments.  We ran similar analyses
for both experiments and found that activity was best reconstructed
when limiting the training data to within task, as opposed to across
task or incorporating data from both tasks (Fig.~\taskrecon (mean reconstruction accuracy incorporating data within task: 0.55, across task: 0.37, all tasks: .50)). Although reconstruction accuracy in the across task analysis was still better than the volume conductance model alone (paired $t$-test
between $z$-transformed mean correlation coefficients by patient: $t(47) = 5.65, p < 10\textsuperscript{-5}$), these results suggests that having a common tasks for patients may yield better reconstruction accuracy.
